---
title: "TBI"
author: "Edwin Zhu"
date: "May 12, 2020"
output: html_document
---

```{r, echo=FALSE}
# Setting up appropriate tables

library(readxl)
library(ggpubr)
library(tidyr)
library(dplyr)
library(fuzzyjoin)
setwd("C:/Users/edwin/Desktop/Volunteer")
data_path = "C:/Users/edwin/Desktop/Volunteer/CANTBI DTI JHU stats.xlsx"
WMF = read_excel(path=data_path, sheet="WMF")
names(WMF) = as.matrix(WMF[1,])
WMF = WMF[-1,]
WMF[] = lapply(WMF, function(x) as.character(x))
WMF = as.data.frame(WMF)

WMF_skeleton = WMF[1:148,]
WMF_skeleton$type = ifelse(grepl("A$", WMF_skeleton$subject), "baseline", "6_week")
# Change the order for type
# Now baseline is the first box in boxplot
WMF_skeleton$type = factor(WMF_skeleton$type, levels = c("baseline","6_week"))
WMF_brain = WMF[152:299,]
WMF_brain$type = ifelse(grepl("A$", WMF_brain$subject), "baseline", "6_week")
# Change the order for type
# Now baseline is the first box in boxplot
WMF_brain$type = factor(WMF_skeleton$type, levels = c("baseline","6_week"))
WMF_skeleton_ROI = WMF_skeleton[,c(1:50,73)]
WMF_skeleton_Tract = WMF_skeleton[,c(1,52:73)]
WMF_brain_ROI = WMF_skeleton[,c(1:50,73)]
WMF_brain_Tract = WMF_skeleton[,c(1,52:73)]
# Split the dataset into 4 separate datasets
# Skeleton tract and ROI and brain ROI and tract
# Create new column that categorizes each row as either baseline or a 6 week patient

WMF_skeleton_ROIbaseline = WMF_skeleton_ROI[grep("A$", WMF_skeleton_ROI$subject),]
WMF_skeleton_ROI6week = WMF_skeleton_ROI[grep("B$", WMF_skeleton_ROI$subject),]
# Split skeleton ROI table into two tables based on baseline or 6 week

# WMF_skeleton
# WMF_skeleton_baseline[]
# WMF_skeleton_6week[]
```

```{r, echo=FALSE}
# basic use of shapiro test and plotting
# Does not include sex and age
# Does not separate test and base cases

ggplot (WMF_skeleton_ROIbaseline, aes(x=as.numeric(`All ROIs`))) + geom_histogram(color="black", fill="white")
# ggdensity(as.numeric(WMF_skeleton_ROIbaseline$`All ROIs`), main = "test")
shapiro.test(as.numeric(WMF_skeleton_ROIbaseline$`All ROIs`))
# P-value > 0.05 assume normality

ggplot (WMF_skeleton_ROI6week, aes(x=as.numeric(`All ROIs`))) + geom_histogram(color="black", fill="white")
shapiro.test(as.numeric(WMF_skeleton_ROI6week$`All ROIs`))
# Assume normality

t.test((as.numeric(WMF_skeleton_ROIbaseline[,2])), (as.numeric(WMF_skeleton_ROI6week[,2])))

ggplot(WMF_skeleton_ROI, aes(x=as.numeric(`All ROIs`), fill=type, color=type)) +
  geom_histogram(position="identity", alpha=0.5)

ggplot(WMF_skeleton, aes(type, as.numeric(`All ROIs`))) +
  stat_boxplot(geom = "errorbar") + 
  geom_boxplot( aes(type, as.numeric(`All ROIs`))) +    
  xlab("Type") + 
  ylab("All ROIs")

# wilcox.test((as.numeric(WMF_skeleton_baseline[,2])), (as.numeric(WMF_skeleton_6week[,2])))

# histograms run for both baseline and 6 week for the skeleton based on all ROIs
# Both follwed a normal distribution based on the shapiro test
# t.test shows that there is a difference between means 
```

```{r, echo=FALSE}
data_path_labels = "C:/Users/edwin/Desktop/Volunteer/CANTBIANationalBioba_DATA_LABELS_2020-01-13_0952.xlsx"
labels = read_excel(path=data_path_labels)
# Join two different tables based on subject
# Used fuzzyjoin
colnames(labels)[1] = "subject"
labels = labels %>% drop_na(subject)

# Change the order of levels for labels before joining
labels$`CAN-TBI Study ID (Screening/Hospital data)`= factor(labels$`CAN-TBI Study ID (Screening/Hospital data)`, levels = unique(labels$`CAN-TBI Study ID (Screening/Hospital data)`))

WMF_skeleton_ROI_label = regex_inner_join(WMF_skeleton_ROI, labels, by = "subject")
# Create column to differentiate between control and test groups
WMF_skeleton_ROI_label$group = ifelse(grepl("TC{1}", WMF_skeleton_ROI_label$subject.x), "control", "test")

# Remove cases with no second scan at 6 weeks
temp = names(which(table(WMF_skeleton_ROI_label$`Record ID`) == 1))
WMF_skeleton_ROI_label = WMF_skeleton_ROI_label[!(WMF_skeleton_ROI_label$`Record ID` %in% temp),]
WMF_skeleton_ROI_label

# Run every control group through shapiro test to check normality
shapiro_control_index_norm = vector("numeric")
shapiro_control_index_nonnorm = vector("numeric")
j = 0
k = 0
for (i in 2:50){
  result = shapiro.test(as.numeric(WMF_skeleton_ROI_label[94:148,i]))
  if(result$p.value>0.05){
    shapiro_control_index_norm[j] = i
    j = j + 1
  } else {
    shapiro_control_index_nonnorm[k] = i
    k = k + 1
  }
}

# A few did not follow normal distribution
# Store which columns follow normality in variable, Check using unpaired t test
# Look at normal distrbution using paired t test 
for (i in shapiro_control_index_norm){
  control_normpvalue = t.test(as.numeric(WMF_skeleton_ROI_label[which(WMF_skeleton_ROI_label$type == "baseline" & WMF_skeleton_ROI_label$group == "control"),i]),
         as.numeric(WMF_skeleton_ROI_label[which(WMF_skeleton_ROI_label$type == "6_week" & WMF_skeleton_ROI_label$group == "control"),i]), 
         paired = TRUE, alternative = "two.sided")
  print(i)
  print(control_normpvalue$p.value)
}

#Use wilcox test to test cases that don't follow normality
for (i in shapiro_control_index_nonnorm){
  control_nonnormpvalue = wilcox.test(as.numeric(WMF_skeleton_ROI_label[which(WMF_skeleton_ROI_label$type == "baseline" & WMF_skeleton_ROI_label$group == "control"),i]), as.numeric(WMF_skeleton_ROI_label[which(WMF_skeleton_ROI_label$type == "6_week" & WMF_skeleton_ROI_label$group == "control"),i]), paired = TRUE, alternative = "two.sided")
  print(i)
  print(control_pvalue$p.value)
}

#Run every test group through shapiro test
shapiro_test_index_norm = vector("numeric")
shapiro_test_index_nonnorm = vector("numeric")
j = 0
k = 0
for (i in 2:50){
  result = shapiro.test(as.numeric(WMF_skeleton_ROI_label[1:93,i]))
  if(result$p.value>0.05){
    shapiro_test_index_norm[j] = i
    j = j + 1
  } else {
    shapiro_test_index_nonnorm[k] = i
  }
}
shapiro_test_index
# Store which columns follow normality in variable
```

```{r, echo = FALSE}

# paired t test for control
# For loop to run through every ROI column
for (i in 2:50){
  test_scores = as.numeric(WMF_skeleton_ROI_label[94:148,i])
  print(ggplot(WMF_skeleton_ROI_label[94:148,], aes(x = type, y = test_scores)) +
    stat_boxplot(geom = "errorbar") +
    geom_boxplot( aes(type, test_scores)) +   
    ggtitle("Paired t-test on Control") +
    xlab("Type") +  
    ylab(colnames(WMF_skeleton_ROI_label)[i]))
}

# Paired t test for test subjects
# For loop to run through every ROI column
for (i in 2:50){
  test_scores = as.numeric(WMF_skeleton_ROI_label[1:93,i])
  print(ggplot(WMF_skeleton_ROI_label[1:93,], aes(x = type, y = test_scores)) +
    stat_boxplot(geom = "errorbar") +
    geom_boxplot( aes(type, test_scores)) +   
    ggtitle("Paired t-test on test") +
    xlab("Type") + 
    ylab(colnames(WMF_skeleton_ROI_label)[i]))
}

# Run anova with gender as covariate for control
# Use for loop to go through every single ROI tract
ggplot(WMF_skeleton_ROI_label[94:148,], aes(x = `Gender (Patient Management)`,
                                            y = as.numeric(`All ROIs`),
                                            color = type)) +
  stat_boxplot(geom = "errorbar") +
  geom_boxplot(aes (`Gender (Patient Management)`, as.numeric(`All ROIs`)))

# Anova with gender as covariate for test group
# Use for loop to go through every single ROI tract
ggplot(WMF_skeleton_ROI_label[1:93,], aes(x = `Gender (Patient Management)`,
                                            y = as.numeric(`All ROIs`),
                                            color = type)) +
  stat_boxplot(geom = "errorbar") +
  geom_boxplot(aes (`Gender (Patient Management)`, as.numeric(`All ROIs`)))

# Looking at age to be a covariant
# Subset based on range of ages? 
# ages 10-20, 21-30 etc or split into two groups
WMF_skeleton_ROI_label[which(as.numeric(WMF_skeleton_ROI_label$`Age at time of injury (please enter) (Patient Management)`) > 25),]
```


```{r, echo=FALSE}
FA = read_excel(path=data_path, sheet="FA")
names(FA) = as.matrix(FA[1,])
FA = FA[-1,]
FA[] = lapply(FA, function(x) type.convert(as.character(x)))
FA

MD = read_excel(path=data_path, sheet="MD")
names(MD) = as.matrix(MD[1,])
MD = MD[-1,]
MD[] = lapply(MD, function(x) type.convert(as.character(x)))
MD

AD = read_excel(path=data_path, sheet="AD")
names(AD) = as.matrix(AD[1,])
AD = AD[-1,]
AD[] = lapply(AD, function(x) type.convert(as.character(x)))
AD

RD = read_excel(path=data_path, sheet="RD")
names(RD) = as.matrix(RD[1,])
RD = RD[-1,]
RD[] = lapply(RD, function(x) type.convert(as.character(x)))
RD
```

# test_matrix = matrix(nrow = 4, ncol = 49)
# rownames(test_matrix) = c("AIC", "P-value", "Adjusted P-value", "Effect Size")
# colnames(test_matrix) = colnames(WMF_Skeleton_ROI[2:50])
# 
# comparison_matrix = matrix(nrow = 6, ncol = 49)
# rownames(comparison_matrix) = c("P-value raw Int", "P-value centered Int", "P-value raw age", "P-value centered age", "P-value raw group", "P-value centered group")
# colnames(comparison_matrix) = colnames(WMF_Skeleton_ROI[2:50])
# 
# for (i in 2:50){
#   output = glm(unlist(WMF_Skeleton_ROI[,i]) ~ age + group, data = WMF_Skeleton_ROI)
#   centered_output = glm(unlist(WMF_Skeleton_ROI_Centered[,j]) ~ `Centered Age` + group, data = WMF_Skeleton_ROI_Centered)
#   mean_test = WMF_Skeleton_ROI[which(WMF_Skeleton_ROI$group == 1), i]
#   mean_test = unlist(mean_test[!is.na(mean_test),1])
#   mean_control = WMF_Skeleton_ROI[which(WMF_Skeleton_ROI$group == 0), i]
#   mean_control = unlist(mean_control[!is.na(mean_control),1])
#   test_matrix[1,i-1] = output$aic
#   test_matrix[2,i-1] = summary(output)$coefficients[3,4]
#   #test_matrix[2,i-1] = mean(mean_test)
#   #test_matrix[3,i-1] = sd(mean_test)
#   #test_matrix[4,i-1] = mean(mean_control)
#   #test_matrix[5,i-1] = sd(mean_control)
#   test_matrix[4,i-1] = (mean(mean_test) - mean(mean_control))/
#                           (sqrt((length(mean_test) - 1) * sd(mean_test)^2 + (length(mean_control) - 1) * sd(mean_control)^2)/
#                                   (length(mean_test) + length(mean_control) -2))
#   if (i == 50){
#     test_matrix[3,] = p.adjust(test_matrix[2,], method = "fdr")
#   }
#   comparison_matrix[1,i-1] = summary(output)$coefficient[1,4]
#   comparison_matrix[2,i-1] = summary(centered_output)$coefficient[1,4]
#   comparison_matrix[3,i-1] = summary(output)$coefficient[2,4]
#   comparison_matrix[4,i-1] = summary(centered_output)$coefficient[2,4]
#   comparison_matrix[5,i-1] = summary(output)$coefficient[3,4]
#   comparison_matrix[6,i-1] = summary(centered_output)$coefficient[3,4]
# }
# test_matrix
# comparison_matrix